---
title: "Final Project"
subtitle: "Data Science for Biologists, Spring 2021"
author: "Holly Felker"
output: 
  html_document:
    theme: spacelab
    highlight: breezedark
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
# DON'T CHANGE THE NEXT LINE, except to add NEW options if you have any.
knitr::opts_chunk$set(warning=FALSE, message=FALSE, echo = TRUE, include = TRUE)

# Load all libraries you use HERE:

library(tidyverse)
library(colorblindr)
```

<br><br>

## Introduction

This dataset is a large compilation of all of the sightings of squirrels within Central Park in New York City over a two week period in 2018. It details their age, coloring, location, and behaviours of over 3000 squirrels. 

I chose this dataset at first because I'd wanted to do something with animals while looking through the past Tidy Tuesdays, and all of the information about the squirrels seemed interesting. Then I actually read through the Squirrel Census website and honestly, it's really charming and cute, and made me want to look into their data that much more. Squirrels in general are adorable and smart (they can remember where they've buried a single nut in a large park for a couple years afterwards, and they will even mimic burying nuts to confuse other "thief" squirrels and birds), and it will be cool to look into a large population dataset of them and see how they generally act.

<br>
You can find the [NYC Squirrel Census data here](https://github.com/rfordatascience/tidytuesday/tree/master/data/2019/2019-10-29). 

(I would recommend also reading the actual [Squirrel Census](https://www.thesquirrelcensus.com/) webpage that the data is from because it's *wild*)

<center>

![*Picture taken from Squirrel Census website*](https://images.squarespace-cdn.com/content/v1/603eb468703ac05daa645759/1615241445862-14IGJV192HTVQU5L40N7/ke17ZwdGBToddI8pDm48kHUV6dDLfCQ4HBz_x2OKkK9Zw-zPPgdn4jUwVcJE1ZvWhcwhEtWJXoshNdA9f1qD7bNalm2hBiow6Ud3tcGmvvHIF2jj_e0-UIP_F6xVGbtXQQjOVi2ID8SIgmDZ4pM6MA/squirrel_haunches_web.png?format=300w)

</center>

```{r read-examine-data}

squirrels <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-10-29/nyc_squirrels.csv")

dplyr::glimpse(squirrels)

```

<br><br>

## Question 1

#### How many squirrels were found in trees compared to on the ground?

### Methods

```{r question_1, fig.width = 6, fig.height = 4}

squirrels %>%
  select(unique_squirrel_id, location) %>% #looking for squirrels and their locations
  distinct() %>% #only getting distinct squirrels and IDs
  drop_na() %>% #getting rid of all unknown locations
  arrange(unique_squirrel_id) %>% #making it easier to look at while checking the tibble
  group_by(location) %>% 
  count(location) %>% #getting the location count
  ungroup() %>% #must ungroup!
ggplot(aes(x = location, y = n)) + #graph time!
  geom_col(aes(fill = location), width = 0.5) + #making a bar graph with the numbers from the counts
  geom_text(aes(label = n), nudge_y = 100) + #it took a bit to remember how to add and fix my labels
  scale_fill_manual(values = c("deepskyblue", "forestgreen")) + #color coding for the different locations
  theme_classic() +
  theme(legend.position = "none") + #don't need the legend for this one
  labs(x = "Squirrel Location", y = "Number of Squirrels") -> ground_plot


#cvd_grid(ground_plot)    checked before finalising that the graph was colorblind-friendly and it looks good!

print(ground_plot) #printing the graph so it shows in the html

```


### Answer

There were many more squirrels found on the ground than there were in the trees. Of the squirrels that had their locations noted, 2111 of them were first seen on the ground, while only 843 of them were seen in trees.


<br><br>

## Question 2

#### Is there a trend of squirrels behaving a certain way around humans depending on their age?

### Methods

```{r question_2}

squirrels %>%
  select(unique_squirrel_id, age, approaches, indifferent, runs_from) %>%
  distinct() %>%
  drop_na() %>%
  group_by(age) %>%
  filter(age != "?") %>%
  pivot_longer(cols = c("approaches", "indifferent", "runs_from"), names_to = "Behaviour", values_to = "value") %>%
  filter(value != "FALSE") %>%
  count(Behaviour)




```


### Answer

Replace this text with your answer to your question.

<br><br>

## Question 3

#### What percentage of squirrels have a cinnamon primary fur color?

### Methods

### Answer

Replace this text with your answer to your question.


<br><br>

## Question 4

#### Where were the most squirrels found over the 2 week period?

### Methods

### Answer

Replace this text with your answer to your question.
