---
title: "Final Project"
subtitle: "Data Science for Biologists, Spring 2021"
author: "Holly Felker"
output: 
  html_document:
    theme: spacelab
    highlight: breezedark
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
# DON'T CHANGE THE NEXT LINE, except to add NEW options if you have any.
knitr::opts_chunk$set(warning=FALSE, message=FALSE, echo = TRUE, include = TRUE)

# Load all libraries you use HERE:

library(tidyverse)
library(colorblindr)
```

<br><br>

## Introduction

This dataset is a large compilation of all of the sightings of squirrels within Central Park in New York City over a two week period in 2018. It details their age, coloring, location, and behaviours of over 3000 squirrels. 

I chose this dataset at first because I'd wanted to do something with animals while looking through the past Tidy Tuesdays, and all of the information about the squirrels seemed interesting. Then I actually read through the Squirrel Census website and honestly, it's really charming and cute, and made me want to look into their data that much more. Squirrels in general are adorable and smart (they can remember where they've buried a single nut in a large park for a couple years afterwards, and they will even mimic burying nuts to confuse other "thief" squirrels and birds), and it will be cool to look into a large population dataset of them and see how they generally act.

<br>
You can find the [NYC Squirrel Census data here](https://github.com/rfordatascience/tidytuesday/tree/master/data/2019/2019-10-29). 

(I would recommend also reading the actual [Squirrel Census](https://www.thesquirrelcensus.com/) webpage that the data is from because it's *wild*)

<center>

![*Picture taken from Squirrel Census website*](https://images.squarespace-cdn.com/content/v1/603eb468703ac05daa645759/1615241445862-14IGJV192HTVQU5L40N7/ke17ZwdGBToddI8pDm48kHUV6dDLfCQ4HBz_x2OKkK9Zw-zPPgdn4jUwVcJE1ZvWhcwhEtWJXoshNdA9f1qD7bNalm2hBiow6Ud3tcGmvvHIF2jj_e0-UIP_F6xVGbtXQQjOVi2ID8SIgmDZ4pM6MA/squirrel_haunches_web.png?format=300w)

</center>

```{r read-examine-data}

squirrels <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-10-29/nyc_squirrels.csv")

dplyr::glimpse(squirrels)

```

<br><br>

## Question 1

#### How many squirrels were found in trees compared to on the ground?

### Methods

```{r question_1, fig.width = 6, fig.height = 4}

squirrels %>%
  select(unique_squirrel_id, location) %>% #looking for squirrels and their locations
  distinct() %>% #only getting distinct squirrels and IDs
  drop_na() %>% #getting rid of all unknown locations
  arrange(unique_squirrel_id) %>% #making it easier to look at while checking the tibble
  group_by(location) %>% 
  count(location) %>% #getting the location count
  ungroup() %>% #must ungroup!
ggplot(aes(x = location, y = n)) + #graph time!
  geom_col(aes(fill = location), width = 0.5) + #making a bar graph with the numbers from the counts
  geom_text(aes(label = n), nudge_y = 100) + #it took a bit to remember how to add and fix my labels
  scale_fill_manual(values = c("deepskyblue", "forestgreen")) + #color coding for the different locations
  theme_classic() +
  theme(legend.position = "none") + #don't need the legend for this one
  labs(x = "Squirrel Location", y = "Number of Squirrels") -> ground_plot


#cvd_grid(ground_plot)    checked before finalising that the graph was colorblind-friendly and it looks good!

print(ground_plot) #printing the graph so it shows in the html

```


### Answer

There were many more squirrels found on the ground than there were in the trees. Of the squirrels that had their locations noted, 2111 of them were first seen on the ground, while only 843 of them were seen in trees.


<br><br>

## Question 2

#### Is there a trend of squirrels behaving a certain way around humans depending on their age?

### Methods

```{r question_2, fig.height = 4, fig.width = 7}

squirrels %>%
  select(unique_squirrel_id, age, approaches, indifferent, runs_from) %>% #getting the necessary columns
  distinct() %>%
  drop_na() %>% #same distinctness and getting rid of NA values
  group_by(age) %>%
  filter(age != "?") %>% #there were some where they didn't know the age of the squirrel, but I got rid of those
  rename(c("Approaches Human" = approaches, 
           "Indifferent towards Human" = indifferent, 
           "Runs from Human" = runs_from)) %>% #renaming all of them so it looks better in the plot
  pivot_longer(cols = c("Approaches Human", 
                        "Indifferent towards Human", 
                        "Runs from Human"), 
               names_to = "behaviour", values_to = "value") %>% #pivoting it so everything is tidier
  filter(value != "FALSE") %>% #only want the true values
  count(behaviour) %>% #should have 6 rows for the 2 ages and 3 behaviours
ggplot(aes(x = age, y = n)) +
  geom_point(aes(fill = age), shape = 23, size = 4) + #getting ~fancy~ with the shapes
  geom_segment(aes(xend = age, y = 0, yend = n)) + #made a lollipop plot for flavor
  geom_text(aes(label = n), nudge_y = 100) + #hey I can actually see the values now
  scale_fill_manual(values = c("deeppink", "gold")) +
  facet_wrap(vars(behaviour)) + #separating them by the different behaviours
  theme_classic() +
  theme(legend.position = "none", 
        strip.background = element_rect(fill = "skyblue")) + #making a nice background strip color
  labs(x = "Squirrel Age", y = "Number of Squirrels") -> behaviour_plot

 
#cvd_grid(behaviour_plot)   passes the colorblind test

print(behaviour_plot)


```


### Answer

It doesn't really seem as though age plays much of a factor in squirrel behaviour. Although the vast majority of squirrels seen were adults, in both age groups, the highest behaviour count seen was that they were indifferent to people nearby. I had previously thought that possibly the juveniles, who wouldn't know any better, might approach the people more because they're curious; or that there would have been a larger number of adult squirrels who approached people, if they had learned that people would give them bits of food while walking around Central Park. But for each behaviour category, it doesn't seem as though age mattered, as, regardless of their age, the majority of all squirrels were indifferent towards humans, while the least amount of squirrels actually approached people.


<br><br>

## Question 3

#### What percentage of squirrels have a cinnamon primary fur color?

### Methods

```{r, question_3}

#mutate, add up everything then divide. Then multiply by 100


squirrels %>%
  select(unique_squirrel_id, primary_fur_color) %>%
  distinct() %>%
  drop_na()



```


### Answer

Replace this text with your answer to your question.


<br><br>

## Question 4

#### Where were the most squirrels found over the 2 week period?

### Methods

```{r, question_4}

#Get the list of all the hectares, then plot the top 10? (or top 5 or something)


```


### Answer

Replace this text with your answer to your question.
